---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---


```{r}
library(readr)
library(ggplot2)
library(hms)
library(dplyr)

rm(list=ls())
```

```{r}
df<- read_csv("sample_orders.csv")

```

```{r}
#Cleanup
str(df)
head(df)

#summary(df$time_opened)
df_2<-df[!is.na(df$time_opened),]
summary(df_2$time_opened)
summary(df_2$customer_id)
summary(df_2$subtotal)
df_3<-df_2[df_2$subtotal!=0,]

df_3$customer_id<-as.factor(df_3$customer_id)

#Remove partial day
df_4<-df_3[as.Date(df_3$time_opened)<'2021-03-21',]

df_f<-df_4
```

```{r}
#Feature Construction
df_f$sale_dol<-df_f$subtotal/100
df_f$dow<-weekdays(as.Date(df_f$time_opened))
df_f$time_only<-as.numeric(as_hms(df_f$time_opened))/3600
```


```{r}
#EDA
ggplot(data=df_f,aes(x=sale_dol))+
         geom_density()+
  ggtitle("Density of Sales in Dollars")

ggplot(data=df_f,aes(x=as.Date(time_opened)))+
         geom_line(stat='count')+
  ggtitle("# Sales by Day")

#table(as.Date(df_f$time_opened))

table(df_f$dow)

ggplot(data=df_f,aes(x=time_only))+
  geom_density()+
  ggtitle("Hour of Day of Orders")

```
# Question 1

The prompt calls for an email style message. For time purposes I will go with bullet points and potentially circle back to an email format time permitting.

Early EDA Insights 
- Majority of sales are in about 15-25 dollar range. See a double peak would like to investigate further. Long tail as expected seems to be almost all sales under 100
- Sales data is from start of Feb to Mar 21. Strong rise of sales over time. Partial last day removed
- Sales Mon-Thur are pretty consistent. Large uptick Fri/Sat/Sun with Sat being pick. Fri/Sun are about 50% higher than average weekday. Sat is about double a normal weekday. (Note should check is we have different number of days of week in dataset, partial weeks, would be better to average by day. Skipping for time purposes.)
- The time of orders comes in two major peaks. About a half hour after noon is the first peak for lunch. About 6:30 is the dinner peak. Would be interesting to see how this changes by day of week.


```{r}
#Feature Engineering by user

#Would probably be more efficient with vectorizing. Im just going to for loop it so I can make faster. Data is small enough

user_list<-unique(df_f$customer_id)

df_user<-data.frame(user_list)
colnames(df_user)<-"user"

df_user$num_orders<-NA
df_user$average_price<-NA
df_user$max_price<-NA
df_user$min_price<-NA
df_user$first_order<-NA
df_user$last_order<-NA
df_user$first_order_day<-NA
df_user$last_order_day<-NA
df_user$prop_order_weekend<-NA

for (i in 1:nrow(df_user)){
  spc_user<-df_user[i,"user"]
  spc_df<-df_f[df_f$customer_id==spc_user,]
  
  df_user$num_orders[i]<-nrow(spc_df)
  df_user$average_price[i]<-mean(spc_df$sale_dol)
  df_user$max_price[i]<-max(spc_df$sale_dol)
  df_user$min_price[i]<-min(spc_df$sale_dol)
  df_user$first_order[i]<-min(spc_df$time_opened)
  df_user$last_order[i]<-max(spc_df$time_opened)
  df_user$first_order_day[i]<-as.Date(min(spc_df$time_opened))
  df_user$last_order_day[i]<-as.Date(max(spc_df$time_opened))
  df_user$prop_order_weekend<-sum(spc_df$dow %in% c("Saturday","Sunday"))/length(spc_df$dow)
}

```

```{r}
ggplot(data=df_user,aes(x=average_price,y=num_orders))+
  geom_point()+
  ggtitle("Number of Orders by Average Order price")

ggplot(data=df_user,aes(as.Date(x=last_order_day,origin='1970-01-01'),y=as.Date(first_order_day,origin='1970-01-01')))+
  geom_point()+
  ggtitle("First/Last Order Date")+
  xlab("Last Order Day")+
  ylab("First Order Day")


```
```{r}
#Splitting into cohorts. I want to see behavior over time and forecast how new users will tracks. I have data for 57 days. Im going to split into 8 cohorts each represented by 1 week of customers whose first order occured in a specific week (7 day period). Anyone who placed an order in the last day will be discarded.

df_user$cohort<-(df_user$first_order_day-18650)%/%7
df_user<-df_user[df_user$cohort!=8,]
```

```{r}
df_user$retention_days<-1+df_user$last_order_day-df_user$first_order_day
df_user$days_data<-max(df_user$last_order_day)-df_user$last_order_day+1



```

```{r}
coh_ind_df<-unique(df_user[,c("user","cohort","first_order_day")])
merge_df<-merge(df_f,coh_ind_df,by.x="customer_id",by.y="user",all.x=TRUE)
merge_df_2<-merge_df[!is.na(merge_df$cohort),]
merge_df_2$order_day<-as.numeric(as.Date(merge_df_2$time_opened))
merge_df_2$days_since_first_order<-merge_df_2$order_day-merge_df_2$first_order_day
```


```{r}
cohorts<-unique(df_user$cohort)
days<-0:max(df_user$retention_days)

coh_df<-expand.grid(cohorts,days)
colnames(coh_df)<-c("cohort","days_since_first_order")

coh_df$ret_count<-NA

coh_df$cum_dol<-NA

for (i in 1:nrow(coh_df)){
  spc_coh<-coh_df[i,"cohort"]
  spc_days_since<-coh_df[i,"days_since_first_order"]
  
  spc_df<-merge_df_2[merge_df_2$cohort==spc_coh & merge_df_2$days_since_first_order>=spc_days_since,]
  coh_df[i,"ret_count"]<-length(unique(spc_df$customer_id))
  
  spc_df_2<-merge_df_2[merge_df_2$cohort==spc_coh & merge_df_2$days_since_first_order<=spc_days_since,]
  coh_df[i,"cum_dol"]<-sum(spc_df_2$sale_dol)
  
}

coh_dol<-merge_df_2 %>%
  group_by(as.factor(cohort)) %>%
  summarise(tot_dol=sum(sale_dol))
colnames(coh_dol)[1]<-"cohort"

coh_tot_df<-table(df_user$cohort)

coh_df_2<-merge(coh_df,coh_tot_df,by.x="cohort",by.y="Var1")
colnames(coh_df_2)[5]<-"total_coh_count"
coh_df_2$prop_ret<-coh_df_2$ret_count/coh_df_2$total_coh_count


coh_df_3<-merge(coh_df_2,coh_dol,by.x="cohort",by.y='cohort')
coh_df_3$prop_dol<-coh_df_3$cum_dol/coh_df_3$tot_dol

```

```{r}
ggplot(data=coh_df_3,aes(x=days_since_first_order,y=prop_ret,color=as.factor(cohort)))+
  geom_point()+
  geom_line()
```
```{r}
ggplot(data=coh_df_3,aes(x=days_since_first_order,y=prop_dol,color=as.factor(cohort)))+
  geom_point()+
  geom_line()
```
2. Cohort Analysis

First a massive pinch of salt. This data is missing in many ways. We only have part of the life of our customers in many ways. Customers who started in the later part of the dataset are effectlivey being censored as we are missing data on their later orders. Because of this many of the later cohors 7,6,5 should will have very different data. Over time I expect the cohots to start to look more and more similar as later data rolls in.

We saw a huge percent of customers are 1 time customers. From the early cohorts we see ~30% cusomters ever place a second order. Dollar wise about 60-70% of money is made just on initial sale with the rest trickling in later.

Because of this I see the second order as a huge oppurtunity. 
